---
title: "ggplot tutorial"
subtitle: "with kunoichi + ninjutsu theme"
author: "<br><br> Emi Tanaka"
date: "<br>2018/09/16"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: "16:9"
      highlightLines: true
---



```{r}
library(tidyverse)
library(readxl)
library(lubridate)
df_0 <- read_xlsx("us_avg_tuition.xlsx") %>% 
  left_join(data_frame(state.abb, State = state.name))
  
library(ggrepel)

ggplot(df_0) +
  aes(x = `2004-05`) +
  aes(y = `2014-15`) +
  labs(x = "Tuition cost in 2004, USD") +
  labs(y = "Tuition cost in 2014, USD") +
  coord_equal() +
  geom_abline(slope = 1, intercept = 0, lty = 2) +
  theme_bw(base_family = "serif") +
  # gain
  geom_segment(
    aes(xend=`2004-05`, yend=`2004-05`, 
        col="Increase from\n2004 to 2014"), alpha = .7) + 
  #original
  geom_segment(aes(xend=`2004-05`, y=0, yend=`2004-05`, col="2004 Tuition Cost" ), alpha = .7) +
  # colscale
  scale_color_manual(breaks = c("Increase from\n2004 to 2014", "2004 Tuition Cost"),values=c("grey70", "grey30")) +
    aes(label = state.abb) +
    geom_point(col = "grey30") +
  lims(x = c(0,15000)) +
  lims(y = c(0,17000)) +
  annotate(geom = "text", x = 1700, y = 1700, 
           label = "Average Tuition\nCosts in 2004", 
           angle = 45, size = 2.8) +
  annotate(geom = "text", x = 13300, y = 15200, 
           label = "Increase", angle = 90, size = 2.8) +
  annotate(geom = "segment", x = 12900, y = 12900, 
           xend = 12900, yend = 17000, 
           arrow = arrow(length = unit(0.25, "cm")), 
           col = "darkgrey") +
  annotate(geom = "segment", 
           x = 12900, y = 12900, xend = 12900, yend = 0, 
           arrow = arrow(ends = "both", length = unit(0.15, "cm"), 
                         angle = 90), col = "darkgrey") +
  annotate(geom = "text", x = 13300, y = 8000, 
           label = "Cost in 2004", angle = 90, size = 2.8) +
  ggpmisc::stat_dens2d_filter(geom = "text_repel", 
                              keep.fraction = 0.30, 
                              size = 3.8) +
  labs(subtitle = "Vis: Gina Reynolds | Data Source: onlinembapage.com") +

  labs(title = "Average in-state tuition fees at public 4-year: 2004 and 2014") +
  labs(col = "") 

```
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(fig.height=6, out.width="100%")
library(ggplot2)
# reveal lines up to `upto` and highlight lines `highlight`
reveal <- function(name, upto, highlight = upto) {
  content <- knitr:::knit_code$get(name)
  content[upto] <- gsub("+", "", content[upto], fixed=T)
  content[highlight] <- paste(content[highlight], "#<<")
  content[1:upto]
}

partial_knit_chunks <- function(chunk_name) {
  # Create slide for lines 1:N for each line N in the given chunk
  idx_lines <- seq_along(knitr:::knit_code$get(chunk_name))
  partial_knit_steps <- glue::glue(
    "class: split-40",
    "count: false",
    "",
    ".column[.content[",
    "```{r plot{{idx_lines}}, eval=FALSE, code=reveal('{{chunk_name}}', {{idx_lines}})}",
    "```",
    "]]",
    ".column[.content.center[",
    "```{r output{{idx_lines}}, echo=FALSE, code=reveal('{{chunk_name}}', {{idx_lines}})}",
    "```",
    "]]",
    .open = "{{", .close = "}}", .sep = "\n"
  )
  glue::glue_collapse(partial_knit_steps, "\n---\n")
}
```

```{r complete, eval=F, echo=F}
ggplot(df_0) +
  aes(x = `2004-05`) +
  aes(y = `2014-15`) +
  coord_equal() +
  labs(x = "Tuition cost in 2004, USD") +
  labs(y = "Tuition cost in 2014, USD") +
  geom_segment(aes(xend=`2004-05`, yend=`2004-05`, col="Increase from\n2004 to 2014"), alpha = .7) + 
  geom_segment(aes(xend=`2004-05`, y=0, yend=`2004-05`, col="2004 Tuition Cost" ), alpha = .7) +
  scale_color_manual(breaks = c("Increase from\n2004 to 2014", "2004 Tuition Cost"),values=c("grey70", "grey30")) +
 aes(label = state.abb) +
  geom_point(col = "grey30") +
  ggpmisc::stat_dens2d_filter(geom = "text_repel", keep.fraction = 0.30, size = 3.8) +
  lims(x = c(0,15000)) +
  lims(y = c(0,17000)) +
  annotate(geom = "text", x = 1700, y = 1700, label = "Average Tuition\nCosts in 2004", angle = 45, size = 2.8) +
  annotate(geom = "text", x = 13300, y = 15200, label = "Increase", angle = 90, size = 2.8) +
  annotate(geom = "segment", x = 12900, y = 12900, xend = 12900, yend = 17000, arrow = arrow(length = unit(0.25, "cm")), col = "darkgrey") +
  annotate(geom = "segment", x = 12900, y = 12900, xend = 12900, yend = 0, arrow = arrow(ends = "both", length = unit(0.15, "cm"), angle = 90), col = "darkgrey") +
  annotate(geom = "text", x = 13300, y = 8000, label = "Cost in 2004", angle = 90, size = 2.8) +
  labs(title = "Average in-state tuition fees at public 4-year: 2004 and 2014") +
  labs(subtitle = "Vis: Gina Reynolds | Data Source: onlinembapage.com") +
  labs(col = "") +
    geom_abline(slope = 1, intercept = 0, lty = 2) +
    theme_bw(base_family = "serif") 
```

`r paste(knitr::knit(text = partial_knit_chunks("complete")), collapse = "\n")`

<!--
Create slides for the "complete" chunk
...it's just R Markdown as a vector of strings
...and why not be super concise about it?
-->

<!-- This css chunk can't be before the first class: split-40... lines -->
```{css, eval=TRUE}
.remark-code{ line-height: 2; }
```